Quantum H theorem, variational entropy, and relaxation processes

### 2020-09-11 notasparaarticulo.tex
Se fusionó el archivo que se trabajó con el Dr. Carrillo con el que tiene el trabajo de la obtención de la ecuación de transporte, no se ha modificado (y se debe modificar) el abstract, la intro y las conclusiones

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amssymb}
\usepackage{amsmath}

\title{Quantum H theorem, Variational Entropy, and Relaxation Processes}
\author{labfqot}
\date{March 2020}

@@

\section{Abstract}
We review in detail the quantum formulation of the Boltzamnns' H theorem. Starting from a H functional defined in the energy space, by means a variational procedure, we demonstrate the analogous of the quantum H theorem. At difference of the Tolmans' procedure to demonstrate the H theorem,  we start from the hypothesis of an out of equilibrium, spatially non homogeneous system, and show that when $\frac{dH}{dt}=0$ the system reaches the equilibrium condition and then the proposed H functional becomes the well known expression of the entropy of an ideal quantum gas. The time evolution of the system occurs in such a way that $\frac{dH}{dt}<0$. Based on this, we propose a theoretical scheme that allows to describe the time evolution toward the equilibrium of an ideal quantum gas, it is the relaxation process of an out equilibrium.

@@

When the system is in equilibrium, the quantities $\bar{\mathcal{N}}_M$ and $\bar{\mathcal{E}}_M$ will not depend on the cell number. This is

@@

Inspired by the Hamilton principle, H will be maximized with those restrictions to obtain the equilibrium distribution functions for quantum gasses using the Lagrange multipliers method obtaining

@@

where $\alpha_M(t), \beta_M(t)$ are the Lagrange multipliers. In equilibrium, the energy and the Lagrange multiplier do not depend on the cell number and time, therefore

@@

It is necessary to remark that the microcanonical ensemble is used, but the Bose-Einstein and Fermi-Dirac distribution functions are obtained from the grant canonical ensemble. This result is a consequence of describing the system by its cells. The appearing of the chemical potential is a consequence of this description. Equiprobability assumption from the microcanonical ensemble is recovered because of the chemical potential and the temperature are equal and constants in equilibrium, then the total distribution function is a constant.\\    
From here, one can calculate the entropy from an ideal system, in other words, if one substitute (\ref{distributionequilibrium}) to (\ref{entropy}) one obtains
      H(t)=\sum_M \sum_n \int_V \left[\left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right)\ln \left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}} \right) \right]\mathrm{d}V \pm \int_V \ln \left[\prod_{M} \prod_{n}\left(1 \mp \frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right) \right]\mathrm{d}V\label{H-entropy}.
  This result lets us calculate the entropy of a quantum system with inhomogenities, this functional could be used as a thermodynamic variable to describe quantum systems.
\section{The change of H for relaxation processes}

The functional H has the following form
\begin{equation}
    H(t)=\sum_{M} \int_V\sum_{n} \left[ f_{Mn} \ln f_{Mn} \pm (1 \mp f_{Mn}) \ln (1 \mp f_{Mn}) \right]\mathrm{d}V \label{entropy2},
\end{equation}
where, for simplicity, we have defined $f(\bar{r},\epsilon_{n},t)\equiv f_{Mn}$.
These functional has the following microcanonical restrictions
\begin{eqnarray}
        \int_V\sum_{n}f_{Mn}\mathrm{d}V=\bar{\mathcal{N}}+\Delta_M(t); \ \ \ \ \int_V\sum_{n}\epsilon_{n}f_{Mn}\mathrm{d}V=\bar{\mathcal{E}}+ \delta_M(t), \label{restrictionoutside}
  \end{eqnarray}{}
  where $\bar \mathcal{ N}$ and $\bar \mathcal{E}$ are the mean particle number and the mean energy by cell in equilibrium
  \begin{equation}
      \bar{\mathcal{N}}=\int_V \sum_n \bar{f}_n; \ \ \ \ \bar{\mathcal{E}}=\int_V \sum_n \epsilon_n\bar{f}_n
  \end{equation}
  and $\Delta_M,\delta_M$ could be seen as a deviation from the mean number of particles and mean energy by cell from the equilibrium respectively, with $\Delta_M(t)\ll N$ and $\delta_M(t) \ll E$. Those condition described a system out of equilibrium but no so far from it.  \\
If one performs the derivative of the functional H with respect to time
\begin{equation}
   \frac{dH(t)}{dt}= \sum_n \sum_M \int_{V}\dot{f}_{nM}(t)\ln \left[ \frac{f_{nM}(t)}{1\mp f_{nM}(t)} \right]\mathrm{d}V.\label{deltaH}
\end{equation}{}
Using the first-order approximation, that is
\begin{eqnarray}
   f_{Mn}=\bar{f}_{n}(1+g_{Mn}): \ \ \ \bar{f}_{n}\gg \bar{f}_{n}g_{Mn}, \label{firstorder}
\end{eqnarray}{}
one obtains

@@

Here the deviation $g_{Mn}=g_{Mn}(\bar{r},\epsilon_{n},t)$.\\

@@

   1-\bar f_n -\bar f_n g_{nM}>0 \ \ \Rightarrow \ \ \frac{1}{\bar f_n}>1+g_{nM}.

@@

The previous expression establishes that the value of $g_{nM}$ is determined by $\bar f_n$. Furthermore, whether $\bar f_{n}=0$, this is, all energy levels are occupied, then the system will not have inhomogenities because of the exclusion principle and as consequence of this,  $g_{nM}=0$.\\
In the case of bosons, if $\bar{f}_n \gg \bar{f}_n |g_{nM}|$ then $1+\bar{f}_n \gg \bar{f}_n |g_{nM}|$. In fermions, in the case of a non-extremely degenerated condition, it is true that $1-\bar{f}_n \gg \bar{f}_n |g_{nM}|$. With those relations, one can approximate the logarithm functions to their first-order Taylor series. Besides, this approximation is for both signs for $g_{nM}$.
That approximation applied to logarithm functions in (\ref{cambioH1}) around $\bar f_n g_{nM}=0$ is

@@

with the previous approximation, (\ref{cambioH1}) will be

@@

   \frac{dH}{dt}&=& \bar{\alpha}\sum_M \dot{\Delta}_M(t)+\bar{\beta}\sum_M \dot{\delta}_M(t)+\sum_n \sum_M \int_V\left \{ \bar{f}_{n}g_{nM}\left(1\pm e^{\bar{\alpha}+\bar{\beta}{\epsilon}_n}\right)\dot{g}_{nM} \right \}\mathrm{d}V \nonumber \\
   &=&\sum_n  \bar{f}_{n}\left(1\pm e^{\bar{\alpha}+\bar{\beta}{\epsilon}_n}\right)\sum_M \int_V g_{nM}\dot{g}_{nM}\mathrm{d}V= \sum_n \left(\frac{1\pm e^{\bar{\alpha}+\bar{\beta}\epsilon_n}}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_n}\pm 1} \right)\sum_M \int_V g_{nM}\dot{g}_{nM}\mathrm{d}V. \nonumber \\ \label{cambioH4}

@@

On the one hand, $\dot{g}^{+}_{nJ}<0$ as a consequence of lost of particles or energy and so , $g^{+}_{nJ}>0$. On the other hand, $\dot{g}^{-}_{nJ}>0$ as a consequence of gain of particles or energy and therefore $g^{-}_{nJ}<0$. As a result of the previous idea

@@

    \frac{dH}{dt}=-\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L}\int_V |g^{+}_{nJ}||\dot{g}^{+}_{nJ}|\mathrm{d}V+\sum^{N}_J \int_V |g^{-}_{nJ}||\dot{g}^{-}_{nJ}|\mathrm{d}V \right). \label{cambioH6}

@@

In conclusion, $\frac{dH}{dt}<0$.\\
When the system is in equilibrium, $g_{nM}=0$ and in consequence of that, $\dot g_{nM}=0$ and finally from (\ref{cambioH1}) that $\frac{dH}{dt}=0$.
with the previous prove, in analogous to the classical scheme, the homogeneous distribution function hypothesis must be eliminated to be consistent with this prove. Now, in the following section, variational procedure and the relaxation processes will be used in the classical scheme to prove the $H$ theorem eliminating the homogeneous distribution function hypothesis.

\section{Classical scheme}
\subsection{Variational procedure}
We proposed the following $H$ functional

\begin{equation}
   H(t)=\int_{}^{}\int_{}^{} f(r,v,t) \ln f(r,v,t)\mathrm{d}^3v \mathrm{d}^3r \label{CH2}.
\end{equation}{}
The functional (\ref{CH2}) contains a no-homogeneous distribution function but keeps the same form of the Boltzman's $H$ functional. Excluding the homogeneous distribution function hypothesis, one uses the same procedure, that it was shown in the previous section, to prove the $H$-theorem for relaxation processes.\\ 
If one considers the microcanonical ensemble conditions 
\begin{equation}
    \int_{}^{}\int_{}^{}f(r,v,t)\mathrm{d}^3v \mathrm{d}^3r=N, \ \ \ \int_{}^{}\int_{}^{}f(r,v,t)\epsilon(v)\mathrm{d}^3v \mathrm{d}^3r=E \label{micro},
\end{equation}{}
one obtains the distribution function that minimizes $H$
\begin{eqnarray}
\frac{\delta H}{\delta f(r',v')}&=&\int_{}^{}\frac{\delta}{\delta f(r',v')}\left[f(r,v)\ln f(r,v)  \right]\mathrm{d}^3v \mathrm{d}^3r-\int_{}^{}\alpha(r)\int_{}^{}\frac{\delta f(r,v)}{\delta f(r',v')}\mathrm{d}^3v \mathrm{d}^3r\nonumber \\
&&-\int_{}^{}\beta(r) \int_{}^{}\epsilon(v)\frac{\delta f(r,v)}{\delta f(r',v')}\mathrm{d}^3v \mathrm{d}^3r\nonumber \\
&=&\ln f(r',v')+1-\alpha(r')-\beta(r') \epsilon(v')=0.
\end{eqnarray}{}
isolating the distribution function 
\begin{eqnarray}
\ln f(r',v')&=&\alpha(r')+\beta(r') \epsilon(v')-1 \ \ \  \Rightarrow \ \ \ f(r',v')=e^{\alpha(r') +\beta(r') \epsilon(v')-1} \nonumber \\
&=&Ce^{\alpha(r')+\beta(r') \epsilon(v') } \label{relacion1},
\end{eqnarray}{}
where $C$ is a constant. It is trivial that using the variational procedures, one obtains the form of the Maxwell-Boltzmann distribution. \\
As $H$ in equilibrium is a density of entropy, it is trivial to think that Lagrange multipliers do not depend on the position because of the system is in equilibrium and consequently, the distribution function is homogeneous, that is
\begin{equation}
    f(v)=Ce^{\alpha+\beta \epsilon(v)}.
\end{equation}{}

\subsubsection{Relaxation processes}
Now one uses the restrictions (\ref{restrictionoutside}) but in the continuous case
\begin{equation}
    \int_{}^{}f(r,v,t)\mathrm{d}^3v=\mathcal{N}+\Delta(r,t); \ \ \ \int_{}^{}f(r,v,t)\epsilon(v)\mathrm{d}^3v=\mathcal{E}+\Omega(r,t)\label{restriccionescambio},
\end{equation}{}
where $\Delta(r,t)$ y $\Omega(r,t)$ could be seen as a deviation from the mean number of particles and mean energy by differential element $d^3r$.
Performing the derivative of the functional $H$ (\ref{CH2})with respect to time
\begin{equation}
    \frac{dH}{dt}=\int_{}^{}\int_{}^{}\left[ 1+\ln f(r,v,t) \right]\dot f(r,v,t) \mathrm{d}^3v \mathrm{d}^3r \label{dH1}.
\end{equation}{}
Applying the first order approximation (\ref{firstorder}) to (\ref{dH1}), one obtains
\begin{equation}
    \frac{dH}{dt}=\int_{}^{}\int_{}^{}\bar f(v) \left [ 1+\ln \left\{ \bar f(v)+\bar f(v)g(r,v,t) \right\} \right]\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r \label{dH1,1}.
\end{equation}{}
Using the same approximation (\ref{lnapproximation})
to (\ref{dH1,1}) 
\begin{eqnarray}
\frac{dH}{dt}&=&\int_{}^{} \int_{}^{} \bar f(v)\left[ 1+\ln \bar f(v)+g(r,v,t) \right]\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r,
\end{eqnarray}{}
using (\ref{relacion1}) with homogeneous multipliers
\begin{eqnarray}
\frac{dH}{dt}&=&\int_{}^{}\int_{}^{}\bar f(v)\left[ \alpha+\beta \epsilon(v) \right]\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r+\int_{}^{}\int_{}^{}\bar f(v)g(r,v,t)\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r \label{dH1,2}. \nonumber \\
\end{eqnarray}{}
If one uses the same procedure shown in the previous section to the restrictions (\ref{restriccionescambio}), one obtains the same result. Then, applying that result to (\ref{dH1,2}), one finds
\begin{eqnarray}
\frac{dH}{dt}&=&\int_{}^{}\int_{}^{}\bar f(v)g(r,v,t)\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r.
\end{eqnarray}{}
Using the same consideration in the previous section to $g$ and $\dot g$, one proves that $\frac{dH}{dt}<0$. If the system is in equilibrium, $g(r,v,t)=\dot g(r,v,t)=0$, therefore $\frac{dH}{dt}=0$.

\section{Discussion}
For quantum ideal gases, the H-theorem has been proved performing the variational method. In the classical scheme, the $H$-theorem is proved using the Boltzmann transport equation for a homogeneous distribution function. Classical $H$-theorem don't show the idea of particles transport in a system. That assumption in quantum mechanics presented by Tolman \cite{tolman} remains. This work eliminates that assumption defining the $H$ functional (\ref{entropy}) including a non-homogeneous distribution function. The prove of the quantum $H$-theorem in this work includes the idea of particles transport in relaxation processes using the first order approximation. This idea has not found in the literature. As consequence of this, all ideal gases placed in non-equilibrium state with inhomogeneities, when time tends to infinity, the system tends to the equilibrium state. This result complements the classical $H$-theorem due to the lack of the non-homogeneous distribution function. In addition, the classical $H$ theorem was analysed to include the particles transport and complete the classical $H$-theorem modifying the classical $H$ functional proposed by Boltzmann. Those proves only are applied to relaxation processes but the fact of including a non-homogeneous distribution function is a new way to complete the $H$ theorem for non-equilibrium systems. 
In other hand, the equation (\ref{H-entropy}) lets us calculate the entropy for a quantum system and describe the evolution of that system using the $H$ functional as a thermodynamic variable for a non-equilibrium system. Furthermore, thermodynamic variables to describe non-equilibrium systems is a problem in discussion. The previous idea adds new knowledge to this problem.

\begin{thebibliography}{}
\bibitem{goldstein}
H. Goldstein, {\it Mec\'anica Cl\'asica}, Ed. 1987, Espa\~na: Editorial Revert\'e (2002).

\bibitem{sakurai}
J. J. Sakurai, J. Napolitano {\it Modern Quantum Mechanics}, Second Edition, Addison-Wesley, 1994.

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{xcolor}

\title{Quantum H theorem, variational entropy, and relaxation processes}
\author{labfqot}
\date{March 2020}

@@

\section{Abstract}
We review in detail the quantum formulation of the Boltzamnns' H theorem. Starting from a H functional defined in the energy space. By means a variational procedure, we demonstrate the analogous of the quantum H theorem. At difference of the Tolmans' procedure to demonstrate the H theorem,  we start from the hypothesis of an out of equilibrium, spatially non homogeneous, system, and show that when $\frac{dH}{dt}=0$, the system reaches the equilibrium condition and then proposed H functional becomes the well known expression of the entropy of an ideal quantum gas. The time evolution of the system happens in such a way that $\frac{dH}{dt}<0$. Based on this, we propose a theoretical scheme that allows to describe the time evolution toward the equilibrium of an ideal quantum gas, it is the relaxation process of an out equilibrium.

@@

When the system is in equilibrium, the local number of particles and the local energy don't depend on the cell number. This is

@@

Inspired by the Hamilton principle, $H$ will be maximized with those restrictions to obtain the equilibrium distribution functions for quantum gasses using the Lagrange multipliers method obtaining

@@

where $\alpha_M(t), \beta_M(t)$ are the Lagrange multipliers. \textcolor{red}{We can obtain from the variational procedure that 
    \alpha_M\propto \frac{\partial \ln \Omega_M}{\partial \mathcal{N}_M}, \ \ \ \beta_M\propto \frac{\partial \ln \Omega_M}{\partial \mathcal{E}_M},\label{multipliers}
where we have used that the $H$ functional in equilibrium is proportional to the entropy of the system.
In other hand, in the expression (\ref{multipliers}), we expect that the variation of the logarithm with the number of particles by cells is indifferent from a particular cell. With this, we conclude  
    \alpha_M=\alpha; \ \ \ \ \beta_M=\beta.
\end{equation}}
In equilibrium, the energy and the Lagrange multiplier do not depend on the cell number and time, therefore

@@

We have proved that when the system is in equilibrium, the Lagrange multipliers don't depend on the cell number. In other words, Lagrange multiplier are homogeneous in the system.\\
We can calculate the entropy from an ideal system, in other words, if one substitute (\ref{distributionequilibrium}) to (\ref{entropy}) one obtains
      H(t)&=&\sum_M \sum_n \int_V \left[\left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right)\ln \left(\frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}} \right) \right]\mathrm{d}V\nonumber \\
      &&\pm \int_V \ln \left[\prod_{M} \prod_{n}\left(1 \mp \frac{1}{e^{-\bar{\alpha}-\bar{\beta}\epsilon_{n}}\pm 1} \right) \right]\mathrm{d}V\label{H-entropy}.
  This result lets us calculate the entropy of a quantum ideal gases. This entropy could be used as a thermodynamic variable to describe quantum ideal gases.
\subsection{Classical scheme: variational procedure}
We proposed the following $H$ functional

\begin{equation}
   H(t)=\int_{}^{}\int_{}^{} f(r,v,t) \ln f(r,v,t)\mathrm{d}^3v \mathrm{d}^3r \label{CH2}.
\end{equation}{}
The functional (\ref{CH2}) contains a no-homogeneous distribution function but keeps the same form of the Boltzman's $H$ functional. Excluding the homogeneous distribution function hypothesis, one uses the same procedure, that it was shown in the previous section, to prove the $H$-theorem for relaxation processes.\\ 
If one considers the microcanonical ensemble conditions 
\begin{equation}
    \int_{}^{}\int_{}^{}f(r,v,t)\mathrm{d}^3v \mathrm{d}^3r=N, \ \ \ \int_{}^{}\int_{}^{}f(r,v,t)\epsilon(v)\mathrm{d}^3v \mathrm{d}^3r=E \label{micro},
\end{equation}{}
one obtains the distribution function that minimizes $H$
\begin{eqnarray}
\frac{\delta H}{\delta f(r',v')}&=&\int_{}^{}\frac{\delta}{\delta f(r',v')}\left[f(r,v)\ln f(r,v)  \right]\mathrm{d}^3v \mathrm{d}^3r-\int_{}^{}\alpha(r)\int_{}^{}\frac{\delta f(r,v)}{\delta f(r',v')}\mathrm{d}^3v \mathrm{d}^3r\nonumber \\
&&-\int_{}^{}\beta(r) \int_{}^{}\epsilon(v)\frac{\delta f(r,v)}{\delta f(r',v')}\mathrm{d}^3v \mathrm{d}^3r\nonumber \\
&=&\ln f(r',v')+1-\alpha(r')-\beta(r') \epsilon(v')=0.
\end{eqnarray}{}
isolating the distribution function 
\begin{eqnarray}
\ln f(r',v')&=&\alpha(r')+\beta(r') \epsilon(v')-1 \ \ \  \Rightarrow \ \ \ f(r',v')=e^{\alpha(r') +\beta(r') \epsilon(v')-1} \nonumber \\
&=&Ce^{\alpha(r')+\beta(r') \epsilon(v') } \label{relacion1},
\end{eqnarray}{}
where $C$ is a constant. It is trivial that using the variational procedures, one obtains the form of the Maxwell-Boltzmann distribution. \\
As $H$ in equilibrium is a density of entropy, it is trivial to think that Lagrange multipliers do not depend on the position because of the system is in equilibrium and consequently, the distribution function is homogeneous, that is
\begin{equation}
    f(v)=Ce^{\alpha+\beta \epsilon(v)}.
\end{equation}{} 
The constant $C$ can be omitted defining the following $H$ functional
\begin{equation}
   H(t)=\int_{}^{}\int_{}^{} \left[f(r,v,t) \ln f(r,v,t)-f(r,v,t)\right]\mathrm{d}^3v \mathrm{d}^3r \label{CH3},
\end{equation}{}
  but $\int \int f(r,v,t)\mathrm{d}^3v \mathrm{d}^3r=N$ where $N$ is the total particle number. As $N$ is a constant and the functional $H$ has sense when one calculates the change of $H$, this constant can be omitted. 
\section{H for relaxation processes in out of equilibrium systems}

\subsection{Quantum Scheme}

The functional H has the following form
\begin{equation}
    H(t)=\sum_{M} \int_V\sum_{n} \left[ f_{Mn} \ln f_{Mn} \pm (1 \mp f_{Mn}) \ln (1 \mp f_{Mn}) \right]\mathrm{d}V \label{entropy2},
\end{equation}
where, for simplicity, we have defined $f(\bar{r},\epsilon_{n},t)\equiv f_{Mn}$.
These functional has the following microcanonical restrictions
\begin{eqnarray}
        \int_V\sum_{n}f_{Mn}\mathrm{d}V=\bar{\mathcal{N}}+\Delta_M(t); \ \ \ \ \int_V\sum_{n}\epsilon_{n}f_{Mn}\mathrm{d}V=\bar{\mathcal{E}}+ \delta_M(t), \label{restrictionoutside}
  \end{eqnarray}
  where $\bar {\mathcal{N}}$ and $\bar{\mathcal{E}}$ are the local particle number and the local energy in equilibrium
  \begin{equation}
      \bar{\mathcal{N}}=\int_V \sum_n \bar{f}_n; \ \ \ \ \bar{\mathcal{E}}=\int_V \sum_n \epsilon_n\bar{f}_n,
  \end{equation}
  and $\Delta_M,\delta_M$ could be seen as a deviation from $\bar{\mathcal{N}}$ and $\bar{\mathcal{E}}$ respectively, with $\Delta_M(t)\ll \bar{\mathcal{N}}$ and $\delta_M(t) \ll \bar{\mathcal{E}}$. Those condition described a system out of equilibrium but no so far from it.  \\
If one performs the derivative of the functional H with respect to time, we obtain
\begin{equation}
   \frac{dH(t)}{dt}= \sum_n \sum_M \int_{V}\dot{f}_{nM}(t)\ln \left[ \frac{f_{nM}(t)}{1\mp f_{nM}(t)} \right]\mathrm{d}V.\label{deltaH}
\end{equation}{}
Using the first-order approximation
\begin{eqnarray}
   f_{Mn}=\bar{f}_{n}(1+g_{Mn}): \ \ \ \bar{f}_{n}\gg \bar{f}_{n}|g_{Mn}|, \label{firstorder}
\end{eqnarray}{}
we get

@@

@@

   1-\bar f_n -\bar f_n g_{nM}>0 \ \ \Rightarrow \ \ \frac{1}{\bar f_n}>1+g_{nM}. \label{fermionrestriction}

@@

The previous expression establishes that the value of $g_{nM}$ is determined by $\bar f_n$. Furthermore, whether $\bar f_{n}=1$, this is, all lower energy levels are occupied, then the system will not have inhomogenities because of the exclusion principle and as consequence of this,  $g_{nM}=0$. This result corresponds to the system in the zero temperature condition. We can observe that in this condition, (\ref{fermionrestriction}) is violated. Due to this, the zero temperature condition is excluded.\\
We can observe in the case of bosons, if $\bar{f}_n \gg \bar{f}_n |g_{nM}|$ then $1+\bar{f}_n \gg \bar{f}_n |g_{nM}|$. In fermions, in the case of a non-extremely degenerated condition, it is true that $1-\bar{f}_n \gg \bar{f}_n |g_{nM}|$. With those relations, we can approximate the logarithm functions in (\ref{cambioH1}) to their first-order Taylor series around $\bar f_n g_{nM}=0$ 

@@

With the previous approximation, (\ref{cambioH1}) will be

@@

   \frac{dH}{dt}&=&  \sum_n e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\sum_M \int_V g_{nM}\dot{g}_{nM}\mathrm{d}V. \nonumber \\ \label{cambioH4}

@@

On the one hand, $\dot{g}^{+}_{nJ}<0$ describes the lost of particles and/or energy and so , $g^{+}_{nJ}>0$. On the other hand, $\dot{g}^{-}_{nJ}>0$ describes the gain of particles and/or energy and therefore $g^{-}_{nJ}<0$. \\
We sort the previous ideas in the following form

@@

    \frac{dH}{dt}=-\sum_n  e^{\bar{\alpha}+\bar{\beta}\epsilon_n}\left(\sum_J ^{L}\int_V |g^{+}_{nJ}||\dot{g}^{+}_{nJ}|\mathrm{d}V+\sum^{N}_J \int_V |g^{-}_{nJ}||\dot{g}^{-}_{nJ}|\mathrm{d}V \right), \label{cambioH6}

@@

and due to $e^{\bar{\alpha}+\bar{\beta}\epsilon_n}$ is always positive, then $\frac{dH}{dt}<0$.\\
When the system is in equilibrium, $g_{nM}=0$ and in consequence of that, $\dot g_{nM}=0$ and finally from (\ref{cambioH1}) $\frac{dH}{dt}=0$. 
The next step in this model is to obtain the analogous transport equation through variation methods. In order to verify that transport equation can be obtained by applying the variation method, in the next section, the Boltzmann transport equation will be obtained applying the mentioned method to a modified Boltzmann $H$ functional.

\subsection{Clasical scheme}
Now one uses the restrictions (\ref{restrictionoutside}) but in the continuous case
\begin{equation}
    \int_{}^{}f(r,v,t)\mathrm{d}^3v=\mathcal{N}+\Delta(r,t); \ \ \ \int_{}^{}f(r,v,t)\epsilon(v)\mathrm{d}^3v=\mathcal{E}+\Omega(r,t)\label{restriccionescambio},
\end{equation}{}
where $\Delta(r,t)$ and $\Omega(r,t)$ could be seen as a deviation from the mean number of particles and mean energy by differential element $d^3r$.
Performing the derivative of the functional $H$ (\ref{CH2}) with respect to time
\begin{equation}
    \frac{dH}{dt}=\int_{}^{}\int_{}^{}\left[ 1+\ln f(r,v,t) \right]\dot f(r,v,t) \mathrm{d}^3v \mathrm{d}^3r \label{dH1}.
\end{equation}{}
Applying the first order approximation (\ref{firstorder}) to (\ref{dH1}), one obtains
\begin{equation}
    \frac{dH}{dt}=\int_{}^{}\int_{}^{}\bar f(v) \left [ 1+\ln \left\{ \bar f(v)+\bar f(v)g(r,v,t) \right\} \right]\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r \label{dH1,1}.
\end{equation}{}
Using the same approximation (\ref{lnapproximation})
to (\ref{dH1,1}) 
\begin{eqnarray}
\frac{dH}{dt}&=&\int_{}^{} \int_{}^{} \bar f(v)\left[ 1+\ln \bar f(v)+g(r,v,t) \right]\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r,
\end{eqnarray}{}
using (\ref{relacion1}) with homogeneous multipliers
\begin{eqnarray}
\frac{dH}{dt}&=&\int_{}^{}\int_{}^{}\bar f(v)\left[ \alpha+\beta \epsilon(v) \right]\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r+\int_{}^{}\int_{}^{}\bar f(v)g(r,v,t)\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r \label{dH1,2}. \nonumber \\
\end{eqnarray}{}
If one uses the same procedure shown in the previous section to the restrictions (\ref{restriccionescambio}), one obtains the same result. Then, applying that result to (\ref{dH1,2}), one finds
\begin{eqnarray}
\frac{dH}{dt}&=&\int_{}^{}\int_{}^{}\bar f(v)g(r,v,t)\dot g(r,v,t)\mathrm{d}^3v \mathrm{d}^3r.
\end{eqnarray}{}
Using the same consideration in the previous section to $g$ and $\dot g$, one proves that $\frac{dH}{dt}<0$. If the system is in equilibrium, $g(r,v,t)=\dot g(r,v,t)=0$, therefore $\frac{dH}{dt}=0$.

\section{The Boltzmann transport equation}
\textcolor{red}{The Boltzmann $H$ functional is defined as
\begin{equation}
    H=\int_{}^{}f(\vec{v},t)\ln f(\vec{v},t)\mathrm{d}^{3}v,
\end{equation}
where $f$ is the homogeneous distribution function of the system and $v$ is the velocity of the particles defined in the $\mu$-space. In order to obtain the Boltzmann transport equation, is necessary to consider that the distribution function is not homogeneous, that is, $f=f(\vec{r},\vec{v},t)$, where $r$ is the position of the particles defined in the $\mu$-space. Because of that, we defined a modified Boltzmann $H$ functional as
\begin{equation}
    H'=\int_{}^{}\int_{}^{}f(\vec{r},\vec{v},t)\ln f(\vec{r},\vec{v},t)\mathrm{d}^{3}v \mathrm{d}^{3}r.
\end{equation}
This equation involves the position dependence. This dependence is necessary to obtain the complete Boltzmann transport equation.\\
Due to $H$ functional in equilibrium corresponds to a density of entropy and also, according to the second law of thermodynamics, the entropy is maximum, then the following relation must be hold in equilibrium
\begin{equation}
    \delta H'=0.\label{variation1}
\end{equation}
Assuming that $\vec{r}=\vec{r}(t)$ and $\vec{v}=\vec{v}(t)$, (\ref{variation1}) casts into
\begin{equation}
    0=\left( 1+\ln f \right)\left( \frac{\partial f}{\partial t}+\dot{\vec{r}}\cdot \nabla f+\dot{\vec{v}}\cdot \nabla_{\vec{v}} f \right).
\end{equation}
The previous relation holds if $(1+\ln f)=0$ or $(\frac{\partial f}{\partial t}+\dot{\vec{r}}\cdot \nabla f+\dot{\vec{v}}\cdot \nabla_{\vec{v}} f)=0$.\\
Assuming that $(1+\ln f)=0$, we find that $f$ is a constant and consequently, $(\frac{\partial f}{\partial t}+\dot{\vec{r}}\cdot \nabla f+\dot{\vec{v}}\cdot \nabla_{\vec{v}} f)=0$. In other hand, if $f$ is not a constant, the Boltzmann transport equation will be
\begin{equation}
    \frac{\partial f}{\partial t}+\dot{\vec{r}}\cdot \nabla f+\dot{\vec{v}}\cdot \nabla_{\vec{v}} f=0.\label{transport}
\end{equation}
The equation (\ref{transport}) is hold only in equilibrium. If the system is out of equilibrium, the variation is not equal to zero. As a consequence of the $H$ theorem, the variation is less to zero. In conclusion, is necessary to include in (\ref{transport}) a term that described this behavior, this term, as in \cite{huang}, will be the collision term and will be expressed in such form as
\begin{equation}
    \frac{\partial f}{\partial t}+\vec{v}\cdot \nabla f+\frac{\vec{F}}{m}\cdot \nabla_{\vec{v}} f=\left( \frac{df}{dt} \right)\biggm |_{coll},\label{classictransport}
\end{equation}
where $\vec{F}$ is a external force and $m$ is the mass of the particles.\\
\\
The equation (\ref{classictransport}) correspond to the Boltzmann transport equation. With this conclusion, in the next section, the analogous transport equation will be obtained applying the variation method to (\ref{entropy2}).}

\section{Analogous transport equation for ideal quantum gases}
\textcolor{red}{Consider that the distribution function $f_{nM}$ is a function of the local number of particles $\mathcal{N}_M$, the local number of particles in a quantum energy state $\mathcal{N}_{M}^{n}$ and the time $t$.
    f=f(\mathcal{N}_M,\mathcal{N}_M^{n},t),
where $\mathcal{N}_{M}^{n}$ is defined as
    \mathcal{N}_{M}^{n}=\int f_{nM}\mathrm{d}V.
Besides, the local number of particles is a function of time $\mathcal{N}_J=\mathcal{N}_J(t)$ and the local number of particles in a quantum energy state is a function of time and the local number of particles $\mathcal{N}_J^{j}=\mathcal{N}_J^{j}(\mathcal{N}_M,t)$. This is due to when a cell transfer particles to another cell, the particles in some defined state energy levels must change.\\
In orden to obtain the analogous transport equation to this system, we apply the variation method to the functional $H$. Due to $H$ is considered in equilibrium as a density of entropy, and therefore a maximum, the variation must be zero.\\
Appliying the variation to the functional $H$, we obtain the following equation of motion
    0=\int_{}^{}\sum_M \sum_n \ln \left[ \frac{f_{nM}}{1\mp f_{nM}} \right] \left[ \frac{\partial f_{nM}}{\partial t}+\frac{\partial f_{nM}}{\partial \mathcal{N}_J}\dot{\mathcal{N}}_J+\frac{\partial f_{nM}}{\partial \mathcal{N}_J^{j}}\left( \frac{\partial \mathcal{N}_J^{j}}{\partial t}+\frac{\partial \mathcal{N}_J^{j}}{\partial \mathcal{N}_L}\dot{\mathcal{N}}_L \right) \right]\mathrm{d}V. \label{transportequation1}
This equation is valid in equilibrium. When the system is out of equilibrium, the variation is different from zero. Due to this, an expression has to be added to (\ref{transportequation1}). This term, in analogy to the Boltzmann transport equation, will be the collision term and will be expressed as $\left( \frac{df_{nM}}{dt} \right)\Bigm|_{coll}$. Moreover, this term contributes to the equation of motion per each cell and cell. Therefore, (\ref{transportequation1}) casts into
    \sum_n \sum_M\left( \frac{df_{nM}}{dt} \right)\Bigm|_{coll}&=&\int_{}^{}\sum_M \sum_n \ln \left[ \frac{f_{nM}}{1\mp f_{nM}} \right] \nonumber \\
    &&\times \left[ \frac{\partial f_{nM}}{\partial t}+\frac{\partial f_{nM}}{\partial \mathcal{N}_J}\dot{\mathcal{N}}_J+\frac{\partial f_{nM}}{\partial \mathcal{N}_J^{j}}\left( \frac{\partial \mathcal{N}_J^{j}}{\partial t}+\frac{\partial \mathcal{N}_J^{j}}{\partial \mathcal{N}_L}\dot{\mathcal{N}}_L \right) \right]\mathrm{d}V.  \nonumber \\ \label{transportequation2}
\end{eqnarray}}
\section{Discussion}
